[Re] Multiple dynamical modes of thalamic relay neurons: rhythmic bursting and intermittent phase-locking

### 2016-08-26 article/detorakis-2016.tex
New comments addressed

Before proceeding to the simulations we introduce a few details regarding 
the source code and the integration method which should be accurate and fast
enough. The implementation is a Python class (Python $3.5.1$), which uses
Numpy (version $1.10.4$), Scipy (version $0.17.0$), Matplotlib
($1.5.1$) and Scikit-image (version $0.12.3$) packages. The numerical
integration is carried out using the \emph{ode} method of Scipy 
\emph{integrate} package. Three different methods were tested in this work 
(\emph{dopri5}, \emph{Adams}, \emph{BDF}, \cite{ascher:1998}). \emph{Dopri5}
is the closest numerical method to the one used by \cite{wang:1994}
(fifth-order adaptive size Runge-Kutta method). The user of the present 
implementation can choose one of the three methods at the stage of class
instantiating. For all three methods we measured spike times, membrane
potential amplitudes and the coefficient of variation using as threshold value
$0\, \Rm{mV}$. We found that spike event times for \emph{dopri5}, \emph{Adams}
and \emph{BDF} coincide and they all have the same $CV \simeq 2.3608$. 
However, the amplitude of membrane potential was different. Figure~\ref{Fig:1}
{\bfseries \sffamily A} shows the membrane potential (in $\Rm{mV}$) produced by
\emph{dopri5} plotted against the membrane potential generated by \emph{Adams}
method. Figure~\ref{Fig:1} {\bfseries \sffamily B} shows the membrane potential
of a simulation using \emph{dopri5} against another using \emph{BDF}. The 
\emph{Adams} method has less differences in amplitude from \emph{dopri5} 
than \emph{BDF}. Therefore, we decided to use the \emph{Adams} method 
since it runs faster than \emph{dopri5}. 
%%
\begin{table}[!htbp]
    \centering
    \begin{tabular}{ccc}
        \thickhline
        \multicolumn{3}{c}{Simulation Time} \\ \thickhline
        Figure & Simulation Time ($\Rm{s}$) & Integration Step ($\Rm{ms}$) \\ 
        \rowcolor{LightGray}
        $1$ & $6$ & $0.05$ \\ \rowcolor{Gray}
        $2$ & $15\times period$, $5$ & $0.05$ \\ \rowcolor{LightGray}
        $3$ & $6$ & $0.05$ \\ \rowcolor{Gray} 
        $4$ & $1.5$, $1$, $1$ & $0.05$ \\ \rowcolor{LightGray} 
        $5$ & $5$ & $0.05$  \\ \rowcolor{Gray} 
        $6$ & $5$ & $0.05$  \\ \rowcolor{LightGray} 
        $7$ & $40$, $20$ & $0.05$  \\ \thickhline
    \end{tabular}
    \caption{{\bfseries \sffamily Simulations Time}}
    \label{Table:2}
\end{table}
%%

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.9\textwidth]{figs/Figure1.pdf}
    \caption{{\bfseries \sffamily Amplitude differences} The membrane potential
    ($\Rm{mV}$) is generated using \emph{dopri5} integration method and is 
    plotted against ({\bfseries \sffamily A}) the one generated by \emph{Adams}
    method and ({\bfseries \sffamily B}) \emph{BDF} method.}
    \label{Fig:1}
\end{figure}

@@

The bursting frequency of the model can be affected by the injected external
current (see Fig~$7$ in \cite{wang:1994}, two bursting modes). To verify that
the present implementation is capable of producing similar results we 
captured data\footnote{Data are available in the accompanying github repository
of the present article.} values for the external current from Fig~$7$ of the 
original article (dots in Fig $7$, pg. $27$ in \cite{wang:1994}) using a 
software called PlotDigitizer\footnote{\url{(http://plotdigitizer.sourceforge.net/})}.
So for every data point (external current amplitude), we ran a simulation and
the bursting frequency of the membrane potential was measured. The results are
illustrated in Figure~\ref{Fig:6}. The black curve shows the frequency in 
$\Rm{Hz}$ and the blue curve shows the period in $\Rm{sec}$. Black dots indicate
the frequency and open circles the corresponding period (present implementation),
respectively. Cyan crosses show the original data taken from \cite{wang:1994} 
and magenta pentagons the corresponding period. It is apparent that the
present implementation is able to reproduce Wang's Fig~$7$, despite the fact
that some of the original points do not match the exactly the ones computed
using the reference implementation code. 
%%
\begin{figure}[!htbp]
    \includegraphics[width=1.0\textwidth, left]{figs/Figure6.pdf}
    \caption{{\bfseries \sffamily Frequency and period versus steady input 
    current.} This figure illustrates how an external current (ranging
    from $0$ to $-2\, \Rm{\mu A/cm^2}$ affects the bursts frequency (bursting
    mode) of the model. Black curve indicates frequency (black dots) in
    $\Rm{Hz}$ and blue curve period (white circle discs) in $\Rm{sec}$. 
    Cyan crosses (frequency) and magenta pentagons (period) indicate original
    data extracted from \cite{wang:1994}.}
    \label{Fig:6}
\end{figure}
%%

>>>>

Before proceeding to the simulations we introduce a few details regarding 
the source code and the integration method which should be accurate and fast
enough. The implementation is a Python class (Python $3.5.1$), which uses
Numpy (version $1.10.4$), Scipy (version $0.17.0$), Matplotlib
($1.5.1$) and Scikit-image (version $0.12.3$) packages. The numerical
integration is carried out using the \emph{ode} method of Scipy 
\emph{integrate} package. Three different methods were tested in this work 
(\emph{dopri5}, \emph{Adams}, \emph{BDF}, \cite{ascher:1998}). \emph{Dopri5}
is the closest numerical method to the one used by \cite{wang:1994}
(fifth-order adaptive size Runge-Kutta method). The user of the present 
implementation can choose one of the three methods at the stage of class
instantiating. For all three methods we measured spike times, membrane
potential amplitudes and the coefficient of variation using as threshold value
$0\, \Rm{mV}$. \emph{Dopri5} and \emph{Adams} methods differ in one spike 
event time whilst \emph{dopri5} and \emph{BDF} have a difference of 37 spike
events. However, all three methods (\emph{dopri5}, \emph{Adams} and \emph{BDF})
have the same $CV \simeq 2.3608$. The most stroking difference found for the
amplitude of membrane potential. Figure~\ref{Fig:1} {\bfseries \sffamily A}
shows the membrane potential (in $\Rm{mV}$) produced by \emph{dopri5} plotted 
against the membrane potential generated by \emph{Adams} method.
Figure~\ref{Fig:1} {\bfseries \sffamily B} shows the membrane potential
of a simulation using \emph{dopri5} against another using \emph{BDF}. The 
\emph{Adams} method has less differences in amplitude from \emph{dopri5} 
than \emph{BDF}. Therefore, we decided to use the \emph{Adams} method 
since it runs faster than \emph{dopri5}. 
%%
\begin{table}[!htbp]
    \centering
    \begin{tabular}{ccc}
        \thickhline
        \multicolumn{3}{c}{Simulated Time} \\ \thickhline
        Figure & Simulation Time ($\Rm{s}$) & Integration Step ($\Rm{ms}$) \\ 
        \rowcolor{LightGray}
        $1$ & $6$ & $0.05$ \\ \rowcolor{Gray}
        $2$ & $15\times \text{period}$, $5$ & $0.05$ \\ \rowcolor{LightGray}
        $3$ & $6$ & $0.05$ \\ \rowcolor{Gray} 
        $4$ & $1.5$, $1$, $1$ & $0.05$ \\ \rowcolor{LightGray} 
        $5$ & $5$ & $0.05$  \\ \rowcolor{Gray} 
        $6$ & $5$ & $0.05$  \\ \rowcolor{LightGray} 
        $7$ & $40$, $20$ & $0.05$  \\ \thickhline
    \end{tabular}
    \caption{{\bfseries \sffamily Simulations Time}}
    \label{Table:2}
\end{table}
%%

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.9\textwidth]{figs/Figure1.pdf}
    \caption{{\bfseries \sffamily Amplitude differences} The membrane potential
    ($\Rm{mV}$) is generated using \emph{dopri5} integration method and is 
    plotted (black dots) against ({\bfseries \sffamily A}) the one generated
    by \emph{Adams} method and ({\bfseries \sffamily B}) \emph{BDF} method.
    The light shaded diagonal area indicates a $2 \, \Rm{mV}$ difference.}
    \label{Fig:1}
\end{figure}

@@

The bursting frequency of the model can be affected by the injected external
current (see Fig~$7$ in \cite{wang:1994}, two bursting modes). To verify that
the present implementation is capable of producing similar results we 
captured data\footnote{Data are available in the accompanying github repository
of the present article.} values for the external current from Fig~$7$ of the 
original article (dots in Fig $7$, pg. $27$ in \cite{wang:1994}) using a 
software called PlotDigitizer\footnote{\url{(http://plotdigitizer.sourceforge.net/})}.
So for every data point (external current amplitude), we ran a simulation and
the bursting frequency of the membrane potential was measured. The results are
illustrated in Figure~\ref{Fig:6}. The black curve shows the frequency in 
$\Rm{Hz}$ and the blue curve shows the period in $\Rm{sec}$. Black dots indicate
the frequency and open circles the corresponding period (present implementation),
respectively. Cyan crosses show the original data taken from \cite{wang:1994} 
and magenta pentagons the corresponding period. It is apparent that the
present implementation is able to reproduce Wang's Fig~$7$, despite the fact
that some of the original points do not match the exactly the ones computed
using the reference implementation code. Furthermore, it is worth mentioning 
that for properly replicate Figure~\ref{Fig:6} we had to run quite long
simulations (see Table~\ref{Table:2}). If the signal is too short the results 
can be quite different from the ones given in \cite{wang:1994}.
%%
\begin{figure}[!htbp]
    \includegraphics[width=1.0\textwidth, left]{figs/Figure6.pdf}
    \caption{{\bfseries \sffamily Frequency and period versus steady input 
    current.} This figure illustrates how an external current (ranging
    from $0$ to $-2\, \Rm{\mu A/cm^2}$ affects the bursts frequency (bursting
    mode) of the model. Black curve indicates frequency (black dots) in
    $\Rm{Hz}$ and blue curve period (white circle discs) in $\Rm{sec}$. 
    Cyan crosses (frequency) and magenta pentagons (period) indicate original
    data extracted from \cite{wang:1994}.}
    \label{Fig:6}
\end{figure}
%%
